---
title: "Homework 4"
author: "Tim Potter"
date: "March 3, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(datasets)
library(dplyr)
library(ggplot2)
chick <- ChickWeight
```

# Questions

1)

```{r}
weight0 <- chick %>%
  group_by(Diet) %>%
  summarize(n=n(), avgWeight = mean(weight), sdWeight = sd(weight))
weight0
```


2)

```{r}
count <- chick %>%
  group_by(Chick) %>%
  mutate(Count = n()) %>%
  ungroup()
complete <- count %>%
  filter(Count==12)
head(complete, 3)
```

3)

```{r}
complete <- complete %>%
  group_by(Chick) %>%
  mutate(weightgain = weight - weight[Time == 0]) %>%
  ungroup()
head(complete, 3)
```

4)
Going by the graph, the chicks on diet 3 gained the most weight by day 21, while the chicks on diet 1 gained the least. The chicks on diet 2 had the largest spread of weights, and the chicks on diet 4 had the tightest spread.
```{r}
complete$Diet <- factor(complete$Diet, c(1, 2, 4, 3))

complete %>%
  filter(Time == 21) %>%
  ggplot(aes(x = Diet, y = weightgain)) + geom_boxplot() + ylim(0, NA)

```

5)
Chick 35 gained the most weight. It was on diet 3. 
Chick 24 gained the least weight. It was on diet 2.
```{r}
fin <- complete %>%
  filter(Time == 21) %>%
  arrange(desc(weightgain))
fin
complete %>%
  filter(Time == 21) %>%
  arrange(weightgain)
fin %>%
  ggplot(aes(x = Chick, y = weightgain, fill = Diet)) + geom_col()
```

6)
```{r}
complete %>%
  filter(Time == 0) %>%
  arrange(weight)

complete %>%
  filter(Time == 0) %>%
  arrange(desc(weight))
```
Chicks 27, 28, 29, 33, 36, and 48 had the lowest starting weight.
Chicks 3, 11, 19, and 23 had the heighest starting weight.
```{r}
complete <- complete %>%
  group_by(Time) %>%
  mutate(avgWeight = sum(weight) / n()) %>%
  ungroup()
complete
complete %>%
  filter(Chick %in% c(27,28,29,33,36,48,3,11,19,23)) %>%
  ggplot(aes(x = Time, y = weight, color = Chick)) + geom_line() + geom_smooth(aes(x = Time, y = avgWeight), show.legend = TRUE)
```

Chicks with the lowest starting weight
```{r}
complete %>%
  filter(Chick %in% c(27,28,29,33,36,48)) %>%
  ggplot(aes(x = Time, y = weight, color = Chick)) + geom_line() + geom_smooth(aes(x = Time, y = avgWeight), show.legend = TRUE)
```

Chicks with the heighest starting weight
```{r}
complete %>%
  filter(Chick %in% c(3,11,19,23)) %>%
  ggplot(aes(x = Time, y = weight, color = Chick)) + geom_line() + geom_smooth(aes(x = Time, y = avgWeight), show.legend = TRUE)
```

Note how all the chicks with the heighest starting weight finished below the average weightgain of all the chicks, while several of the lightest chicks ended with a weight greater than the average.

7)
```{r}
complete %>%
  filter(Diet == 1) %>%
  ggplot(aes(x = Time, y = weight, color = Chick)) + geom_line()
complete %>%
  filter(Diet == 2) %>%
  ggplot(aes(x = Time, y = weight, color = Chick)) + geom_line()
complete %>%
  filter(Diet == 3) %>%
  ggplot(aes(x = Time, y = weight, color = Chick)) + geom_line()
complete %>%
  filter(Diet == 4) %>%
  ggplot(aes(x = Time, y = weight, color = Chick)) + geom_line()
complete %>%
  filter(Time == 21) %>%
  ggplot(aes(x = weight, fill = Diet)) + geom_histogram(binwidth = 20)
complete %>%
  filter(Time == 21) %>%
  ggplot(aes(x = weightgain, fill = Diet)) + geom_histogram(binwidth = 20)
```

These plots don't really give us any more information than what we already had.

```{r}
complete <- complete %>%
  mutate(percentOffAverage = ((weight - avgWeight) / avgWeight) * 100)
complete

complete %>%
  filter(Diet == 1) %>%
  ggplot(aes(x = Time, y = percentOffAverage, color = Chick)) + geom_line() + ylim(-100,100)
complete %>%
  filter(Diet == 2) %>%
  ggplot(aes(x = Time, y = percentOffAverage, color = Chick)) + geom_line() + ylim(-100,100)
complete %>%
  filter(Diet == 3) %>%
  ggplot(aes(x = Time, y = percentOffAverage, color = Chick)) + geom_line() + ylim(-100,100)
complete %>%
  filter(Diet == 4) %>%
  ggplot(aes(x = Time, y = percentOffAverage, color = Chick)) + geom_line() + ylim(-100,100)
```

These plots show the percentage a chick differed from the average weight on a given day, grouped by the diet of the chick. From these we can see that the chicks on diet 1 were pretty consistently below the average, 3 and 4 were consistently above the average, and diet 2 was pretty much even across the board. Of course there could be several other variables that we're not taking into account here, *but*, from the data we have examined, it would be safe to say that diets 3 and 4 are the best for gaining weight.